Human tissue specific markers

### 2017-02-01 manuscript/manuscript.Rnw
add the headers for the new version of the manuscript
do tissue specific genes plot

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{txfonts}
\usepackage{fancyref}
\usepackage{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\usepackage[capitalize]{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\pagestyle{fancy}

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -1, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("xtable")
library("caret")
library("pROC")
library("grid")

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced",
                   "housekeeping_genes_superset"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")
@

@@

\begin{figure}
<<tissue.specificity.shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
invisible(tissue.specific.genes[,type:="not specific"])
invisible(tissue.specific.genes[,housekeeping:="not housekeeping"])
invisible(tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"])
invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                                 possibly.specific==FALSE,housekeeping:="housekeeping"])
invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                          possibly.specific==TRUE,housekeeping:="housekeeping and specific"])

invisible(tissue.specific.genes[,type:=
                          factor(type,
                                 levels=c("not specific","tissue specific",
                                     "housekeeping",
                                     "housekeeping and specific"
                                          ))])
print(ggplot(tissue.specific.genes[gi.max>0],
             aes(y=specificity,x=gi.max,colour=type))+
      ##geom_point()+
      geom_hex(bins=50)+
      ## stat_summary_2d(fun=function(x){length(x)},bins=100)+
      geom_point(aes(colour=housekeeping),data=tissue.specific.genes[housekeeping!="not housekeeping",])+
      scale_x_log10()+
      xlab("Max Expression (FPKM)")+
      ylab("Tissue Specificity")+
      theme(legend.position="top")
    )
@ 
\caption{Shape of tissue specific genes}
\label{fig:tissue_specific_shape}
\end{figure}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
\usepackage[markifdraft,raisemark=0.01\paperheight,draft]{gitinfo2}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\usepackage[capitalize]{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\setlength{\emergencystretch}{20em}
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}
\ExplSyntaxOn
\NewAcroCommand \gene {
  \group_begin:
    \keys_set:nn {acro} {short-format = \itshape}
    \acro_use:n{#1}
  \group_end:
}
\NewAcroCommand \protein { \acro_use:n {#1} }
\ExplSyntaxOff

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -1, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("xtable")
library("caret")
library("pROC")
library("grid")
library("scales")

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced",
                   "housekeeping_genes_superset"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")

tissue.specific.genes[,type:="not specific"]
tissue.specific.genes[,housekeeping:="not housekeeping"]
tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"]
tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                      possibly.specific==FALSE,housekeeping:="housekeeping"]
tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                      possibly.specific==TRUE,housekeeping:="housekeeping and specific"]

tissue.specific.genes[,type:=
                          factor(type,
                                 levels=c("not specific","tissue specific",
                                     "housekeeping",
                                     "housekeeping and specific"
                                          ))]
@

@@

\begin{figure}
<<tissue.specificity.shape.figure,fig.width=4,fig.height=4,echo=FALSE>>=
print(ggplot(tissue.specific.genes[gi.max>0],
             aes(y=specificity,x=gi.max))+
      ##geom_point()+
      geom_hex(bins=50)+
      geom_line(aes(x=x,y=y),inherit.aes=FALSE,
                color="orange",
                data=data.frame(x=c(10,10,tissue.specific.genes[,max(gi.max)]),
                                y=c(1,0.98,0.98)))+
      ## stat_summary_2d(fun=function(x){length(x)},bins=100)+
      geom_point(aes(colour=gsub("And","\nAnd",
                                 capwords(housekeeping))),
                 size=2,
                 data=tissue.specific.genes[housekeeping!="not housekeeping",])+
      scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                    labels = trans_format("log10", math_format(10^.x)))+
      scale_color_brewer("Housekeeping",palette="Set2")+
      scale_fill_continuous("Count")+
      xlab("Max Expression (FPKM)")+
      ylab("Tissue Specificity Index")+
      theme(legend.position=c(1,0),
            legend.justification=c(1,0),
            legend.box.just="right"
            )
    )

@ 
\caption{Tissue specificity index versus maximum expression in one
  tissue. Genes above and to the right of the orange lines are tissue
  specific. Genes which have been previously
  identified~\cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  as housekeeping genes are generally not tissue specific, with the
  exception of %
  %
  \Sexpr{tissue.specific.genes[housekeeping=="housekeeping and specific",gene]}.%
  %
}
\label{fig:tissue_specific_shape}
\end{figure}
